---
title: "earthquake"
author: "HImanshu Pareek"
date: "August 27, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Box plot of EARTHQUAKE data
loading data into R. Filename is "EARTHQUAKE.xlsx"

```{r code1}
filename <- "EARTHQUAKE.xlsx"

library(readxl)
library(dplyr)

earthquakes <- read_excel(filename, na = c("Not Available", "not available", "NOT AVAILABLE", "na", "NA", "-", "_", "null", "NULL"))
head (earthquakes)
```

number of discrepancy is number of rows with na values. This can be calculated by is.na() function

```{r code2}
dataOfNAValues <- earthquakes[rowSums(is.na(earthquakes)) > 0,]
head (dataOfNAValues)
```

Here, dataOfNAValues has 0 rows. It means, there is not any discrepancy in our cata. If we somehow forgot to keep names of possible na strings while reading excel file, we have to use omit function to remove all na or not suitable values.

```{r code3}
earthquakes <- na.omit(earthquakes)
head (earthquakes)
```

We will rename columns to ease of writing

```{r code4}
colNames <- c("time", "lat", "long", "depth", "mag")
colnames(earthquakes) <- colNames
head (earthquakes)
```

Subsetting depth and magnitude columns.

```{r code5}
new_data <- subset(earthquakes, select = c(depth, mag))
head (new_data)
```

Boxplot before removing outliers

```{r code6}
boxplot(new_data$depth, main = "Boxplot for depth in Km (Before removing outliers)")
boxplot(new_data$mag, main = "Boxplot for magnitude (Before removing outliers)")

```

Calculating 5 point summary

```{r code7}
quantileDepth <- quantile(new_data$depth, probs = c(0, 0.25, 0.50, 0.75, 1.0))
quantileMag <- quantile(new_data$mag, probs = c(0, 0.25, 0.50, 0.75, 1.0))
quantileDepth
quantileMag

```

Calculating minimum and maximum of depth and magnitude to be used.

```{r code8}
q1Depth <- quantileDepth[2]
q3Depth <- quantileDepth[4]
q1Mag <- quantileMag[2]
q3Mag <- quantileMag[4]

IQRDepth <- q3Depth - q1Depth
IQRMag <- q3Mag - q1Mag

minDepth <- q1Depth - (1.5 * IQRDepth)
maxDepth <- q3Depth + (1.5 * IQRDepth)
minMag <- q1Mag - (1.5 * IQRMag)
maxMag <- q3Mag + (1.5 * IQRMag)

sprintf("Range of %s should be %f - %f\n", c("Depth", "Magnitude"), c(minDepth, minMag), c(maxDepth, maxMag))

```

Now, we will remove data below minimum limit and above maximum limit.

```{r code9}
new_data <- filter(new_data, depth >= minDepth & depth <= maxDepth & mag >= minMag & mag <= maxMag)
```

Boxplot after removing outliers

```{r code10}
sprintf("After removing outliers")
boxplot(new_data$depth, main = "Boxplot for depth in Km (After removing outliers)")
boxplot(new_data$mag, main = "Boxplot for magnitude (After removing outliers)")
```

Calculating 5 point summary

```{r code11}
quantileDepth <- quantile(new_data$depth, probs = c(0, 0.25, 0.50, 0.75, 1.0))
quantileMag <- quantile(new_data$mag, probs = c(0, 0.25, 0.50, 0.75, 1.0))
quantileDepth
quantileMag

```

We can see that after removing outliers, boxplot is much more symmetric and we don't have any data in outliers so that we can accept all the data in new_data.
