---
title: "automobiles"
author: "HImanshu Pareek"
date: "August 27, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading data into R

```{r code01}
#accessing file
filename <- "AUTOMOBILES.csv"
automobiles <- read.csv(filename, na.strings = c("NA", "na", "not available", "NOT AVAILABLE", "Not Available", "?", "-", "_", "null", "NULL"))
head (automobiles)
```
Data with na values

```{r code02}
naData <- automobiles[rowSums(is.na(automobiles)) > 0, ]
head (naData)
```
Remove na valued data from automobiles

```{r code03}
automobiles <- na.omit(automobiles)
head (automobiles)
```

Change column names for ease of access

```{r code04}
colNames <- c("symboling", "normalized_lossed", "make", "fuel_type", "aspiration",
              "doors", "body_style", "drive_wheels", "engine_loc", "wheel_base",
              "length", "width", "height", "curb_weight", "engine_type", "noc",
              "engine_size", "fuel_system", "bore", "stroke", "comp_ratio", "hp",
              "peak_rpm", "city_mpg", "high_mpg", "price")
colnames(automobiles) <- colNames
head (automobiles)
```

Class of different columns of automobiles data

```{r code05}
colClasses <- sapply(automobiles, class)
colClasses
```

According to NOIR topology, category of data are
symboling - Nominal
normalized_losses - Interval
make - Nominal
fuel_type - Nominal
aspiration - Nominal
doors - Ordinal
body_style - Nominal
drive_wheels - Nominal
engine_loc - Nominal
wheele_base - Interval
length - Interval
width - Interval
height - Interval
curb_weight - Interval
engine_type - Nominal
noc - Ordinal
engine_size - Interval
fuel_system - Nominal
bore - Interval
stroke - Interval
comp_ratio - Rational
hp - Interval
peak_rpm - Interval
city_mpg - Interval
high_mpg - Interval
prive - Interval

We will take the following attributes
1. Nominal - make
2. Ordinal - noc
3. Interval - length
4. Rational - comp_ratio

## 1. Nominal (make)

subsetting data

```{r code06}
make_data <- subset(automobiles, select = make)
head (make_data)
```

grouping data

```{r code07}
library (dplyr)
make_data <- group_by (make_data, make)
head (make_data)
```

making count of each factor

```{r code08}
make_data <- summarize(make_data, count = n())
head (make_data)
```

Pie chart

```{r code09}
pie(make_data$count, labels = make_data$make, main = "Cars sold by companies")
```

Bar plot 

```{r code10}
barplot(make_data$count, names.arg = make_data$make, main = "Bar chart for car companies",
        xlab = "Car companies", ylab = "Number of cars", las = 2, ylim = c(0, 40))
```

Line chart

```{r code11}
plot(make_data$make, make_data$count,
     main = "Line chart for car companies",
     xlab = "Car companies",
     ylab = "Number of cars", las = 2, ylim = c(0, 40))
```
We can see that maximum number of cars has been sold by 'toyota', i.e. mode of this distribution is toyota

## 2. Ordinal (number_of_cylinders)

sbsetting data
```{r code12}
noc_data <- subset(automobiles, select = noc)
head (noc_data)
```

grouping data

```{r code13}
noc_data <- group_by(noc_data, noc)
noc_data <- summarize(noc_data, count = n())
head (noc_data)
```
vector to order the factors

```{r code14}
numOfCyls <- c("one" = 1,
               "two" = 2,
               "three" = 3,
               "four" = 4,
               "five" = 5,
               "six" = 6,
               "seven" = 7,
               "eight" = 8,
               "nine" = 9,
               "ten" = 10
            )
```

extra column to set order of noc

```{r code15}
noc_data$noc <- as.character(noc_data$noc)
noc_data <- mutate(noc_data, lab = numOfCyls[noc])
head (noc_data)
```

sorting w.r.t. lab (noc)

```{r code16}
noc_data <- arrange(noc_data, lab)
noc_data
```

removing lab column from noc_data

```{r code17}
noc_data$noc <- noc_data$lab
noc_data <- subset (noc_data, select = -lab)
noc_data
```

Line plot

```{r code18}
plot(noc_data$noc, noc_data$count, main = "Car numbers v/s #Cylinders",
     xlab = "Number of cylinders", ylab = "Number of cars", las = 1,
     ylim = c(0, 140))
```

Pie chart

```{r code19}
pie(noc_data$count, labels = noc_data$noc,
    main = "Car numbers v/s number of cylinders in the car")
```

Bar plot

```{r code20}
barplot(noc_data$count, names.arg = noc_data$noc, main = "Bar chart for number of cylinders",
        xlab = "Number of cylinder in car", ylab = "Number of cars", las = 1, ylim = c(0, 140))
```
We can see taht maximum number of cars correspond to 4 cylinders in the car, so mode is 4.

## 3. Interval (length)

subsetting data

```{r code21}
length_data <- subset(automobiles, select = length)
head (length_data)
```

Histogram of length of cars

```{r code22}
hist (length_data$length,
      col = "blue",
      main = "Histogram of length of cars",
      ylim = c(0, 40),
      xlab = "Length of car",
      las = 1)
```

Calculate mean, variance and standard deviation

```{r code23}
mean_len <- mean(length_data$length)
var_len <- var(length_data$length)
sd_len <- sd(length_data$length)
sprintf("%s ===> %f", c("Mean", "Variance", "Standard deiation"), c(mean_len, var_len, sd_len))
```

Calculate q-values (5 points)

```{r code24}
q <- quantile(length_data$length, probs = c(0, 0.25, 0.50, 0.75, 1.0))
q
```
We observe that mean is 172.413836, median is 172.40 and mod is 172.50 (approximately)

## 4: Rational - comp_ratio

subsetting data

```{r code25}
comp_data <- subset(automobiles, select = comp_ratio)
head (comp_data)
```
Histogram of compression ratio

```{r code26}
hist (comp_data$comp_ratio,
      col = "blue",
      main = "Histogram of compression ratio",
      ylim = c(0, 150),
      xlim = c(6, 12),
      xlab = "Compression ratio of car",
      las = 1)
```

Another histogram

```{r code27}
hist (comp_data$comp_ratio,
      col = "blue",
      main = "Histogram of compression ratio",
      ylim = c(0, 60),
      xlab = "Compression ratio of car",
      breaks = seq(6, 23, 0.5),
      las = 1)
```
Another hsitogram

```{r code28}
hist (comp_data$comp_ratio,
      col = "blue",
      main = "Histogram of compression ratio",
      ylim = c(0, 60),
      xlim = c(6, 11),
      xlab = "Compression ratio of car",
      breaks = seq(6, 23, 0.5),
      las = 1)
```

Another histogram

```{r code29}
hist (comp_data$comp_ratio,
      col = "blue",
      main = "Histogram of compression ratio",
      ylim = c(0, 60),
      xlim = c(20, 24),
      xlab = "Compression ratio of car",
      breaks = seq(6, 23, 0.5),
      las = 1)
```

Calculate mean, variance and standard deviation

```{r code30}
mean_comp <- mean(comp_data$comp_ratio)
var_comp <- var(comp_data$comp_ratio)
sd_comp <- sd(comp_data$comp_ratio)
sprintf("%s ===> %f", c("Mean", "Variance", "Standard deiation"), c(mean_comp, var_comp, sd_comp))
```

Calculate q-values (5 points)

```{r code31}
q <- quantile(comp_data$comp_ratio, probs = c(0, 0.25, 0.50, 0.75, 1.0))
q
```
We observe that mean is 10.161132, median is 9.0 and mode is 8.75 (approximately)

##Peak RPM

subsetting data
```{r code32}
peak_data <- subset(automobiles, select = peak_rpm)
head (peak_data)
```

Histogram of peak rpm

```{r code33}
hist (peak_data$peak_rpm,
      col = "blue",
      main = "Histogram of peak rpm",
      xlab = "Peak RPM",
      las = 2,
      breaks = seq(4000, 7000, 100))
```

Calculate mean, variance and standard deviation

```{r code34}
mean_pr <- mean(peak_data$peak_rpm)
var_pr <- var(peak_data$peak_rpm)
sd_pr <- sd(peak_data$peak_rpm)
sprintf("%s ===> %f", c("Mean", "Variance", "Standard deviation"), c(mean_pr, var_pr, sd_pr))
```

Calculate q-values (5 points)

```{r code35}
q <- quantile(peak_data$peak_rpm, probs = c(0, 0.25, 0.50, 0.75, 1.0))
q
```

We observe that mean is 5113.836478, median is 5200 and mode is 4750 (approximately)

##City MPG

subsetting data

```{r code36}
city_mpg <- subset (automobiles, select = city_mpg)
head (city_mpg)
```

Histogram of city mpg

```{r code37}
hist (city_mpg$city_mpg,
      col = "blue",
      main = "Histogram of city mpg",
      xlab = "City MPG",
      breaks = seq(15.0, 50.0, 2.5),
      las = 2)
```

Calculate mean, variance and standard deviation

```{r code38}
mean_cm <- mean (city_mpg$city_mpg)
var_cm <- var (city_mpg$city_mpg)
sd_cm <- sd (city_mpg$city_mpg)
sprintf ("%s ===> %f", c("Mean", "Variance", "Standard deviation"),
         c(mean_cm, var_cm, sd_cm))
```

Calculating q-values (5 points)

```{r code39}
q <- quantile (city_mpg$city_mpg, probs = c(0, 0.25, 0.50, 0.75, 1.0))
q
```
We observe that this is left skewed normal distribution for which mean is 26.522013, median is 26 and mod is 23.75 (approximately)
