---
title: "cars"
author: "HImanshu Pareek"
date: "August 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Computing means for given data

Given data is for distances covered by various cars with various speeds.

reading data into R
filename is "CARS.csv"

```{r reading data into R from excel file}
filename <- "CARS.csv"
cars <- read.csv(filename, row.names = 1, na.strings = c("Not Available", "NOT AVAILABLE", "NULL", "NA", "na", "not available", "-", "_", "X"))

head (cars)
```

 
Removing na values (if any)

```{r removing na values (if any)}
cars <- cars[complete.cases(cars), ]
head (cars)
```

 
We will group data with respect to speed of the car

```{r deleting id column}
library(dplyr)
cars <- group_by (cars,speed)
head (cars)
```

 
We will thtan take average of speeds in each group

```{r Creating random indices for sample}
cars <- summarize(cars, mean_dist = mean(dist))
head (cars)
```

 
Now we will calculate weighted means (AM, GM and HM)
 

```{r Select sample from iris data}
weighted_am <- weighted.mean(cars$mean_dist, cars$speed)
weighted_gm <- exp(weighted.mean(log(cars$mean_dist), cars$speed))
weighted_hm <- 1 / (weighted.mean(1 / (cars$mean_dist), cars$speed))

```

printing weighted am, gm and hm
 

```{r none}
sprintf (
    "%s  = %f",
    c("WEIGHTED AM", "WEIGHTED GM", "WEIGHTED HM"),
    c(weighted_am, weighted_gm, weighted_hm)
)
```

 
Here range and observations both are different, so we will consider GM. We can further visualize data using plot

```{r group sample_iris by Species}
plot (
    cars$speed,
    cars$mean_dist,
    main = "Car meandist v/s speed",
    xlab = "Speed of car",
    ylab = "Mean distance"
)
```

 
We can see that relation between mean distance and speed is approximately linear
